A flexible framework to prototype, develop and manage ubiquitous applications for intelligent environments

### 2013-10-05 trunk/ labtempo-ubicomp-projects/UbicompMiddleware/doc/PerCom2014/percom2014_conf.tex
PIPA section rewritten text cleaned to maintain the pages number text finished

The concepts of the framework were materialized on a platform called \textit{SmartAndroid}\footnote{www.tempo.uff.br/smartandroid}, whose development has enabled appraisal in order to prove that the conceptual framework facilitates the process of building UbiApps. We implemented some use cases to validate different aspects of the framework proposed and to testify its capabilities, supported by the elaboration of competency questions, which is a mechanism mostly used to evaluate ontologies, but can also be used as an assessment to this work. 
%power of representation 
%testify the feasibility 
% has enabled the construction of use cases that demonstrate the effectiveness of the approach to provide context awareness to new applications.

% The concepts of the framework have been implemented on a platform called \textit{SmartAndroid}\footnote{www.tempo.uff.br/smartandroid} developed in the context of the project. The implementation of this project has enabled appraisal in order to prove that the conceptual framework facilitates the process of building UbiApps.  One of the evaluations occurred during the processing of an application running on a purely local UbiApp. In another assessment, an application was built to verify the feasibility of implementation of Context Model, through the exploitation of communication mechanisms used in the Distributed Component Model.

The remainder of this paper is organized as follows. First we compare our approach with related works in Section~\ref{sec:related}, then in Section~\ref{sec:overview} we present an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we describe the framework's overall architecture as well as, in Section~\ref{sec:features}, its most important features. Our framework reference implementation as the assumptions, techniques and features related are discussed in Section~\ref{sec:smartandroid}. A proof of concept demonstrating the feasibility of building UbiApps using the framework is presented in Section~\ref{sec:evaluation}, showing examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:conclusion} we conclude this paper with the main remarks and future works.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work} \label{sec:related}
% [ubicomp is nice but it is not priceless]
The benefits to users and developers that UbiComp foresees reach beyond the horizon of many imaginative researchers. The possibilities that arise from it surpass the bounds established by common interface devices, such as mice, keyboards and touch screens. Nonetheless, building such adaptable applications requires much effort from developers who see themselves immersed in a universe of device's specifications and communication technologies, apart from the problem itself that the application must solve. Therefore, many researchers have focused on diminishing those obstacles by providing abstractions, middlewares, services, tools, and other supportive techniques, not only for development but also for prototyping of UbiApps.

@@

%%%%%%%%%%%%%%% CONTEXT AWARENESS
%TODO: Context Toolkit - Dey
%also a famous approach
%we use these abstractions, the aggregation, interpretation, the widget
%This approach provides a starting point
% The Context Toolkit~\cite{dey2001conceptual} is a seminal work that provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation to support rapid prototyping of context aware applications. 
The Dey and Abowd framework~\cite{dey2001conceptual}, implemented by the Context Toolkit, is a seminal work for the supporting of rapid prototyping of context aware applications and provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation. Another pioneer in the field, the Gaia middleware~\cite{ranganathan2003middleware} was designed to facilitate the construction of applications for IE. It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others. However, these two approaches do not have concerns on the user side and their preferences, including the prototyping and management of UbiApps at user level.
% Being one pioneer in the field, 
%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets information, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

% DIOS++
The framework proposed in DIOS++~\cite{liu2003dios++} offers abstractions to manipulate sensors and actuators, a communication layer to manage distributed sensors and actuators, and a distributed rule engine. But concerns on temporization of context rules and creation of rules at end user level were not touched upon.

% CASS
The CASS~\cite{Park2007simulation} is a simulation rule environment that have a simple mechanism to discover possible conflicts between rules that are composed only using ``AND'' clause. However, it does not deploy the rules in a running environment.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview} \label{sec:overview}
% [context/contx sensibility]
It is easy to notice that the context information is a first concern topic in ubiquitous systems. But what is ``context''? Many authors have proposed definitions to this concept, though most are little accurate~\cite{baldauf2007survey}. Synthesizing previous definitions, Dey and Abowd have proposed that context is ``any information that can be used to characterize the situation of entities (i.e., whether a person, place or object) that are considered relevant to the interaction between a user and an application, including the user and the application themselves''~\cite{dey2001conceptual}. In a IE ``Places'' are the rooms, the floors of a building, or spaces in general that enable the localization of other entities, ``People'' are individuals that populate the environment and interact with it, and ``things'' are virtual representations of physical objects or software components.
%Considering the diversity of devices, systems and protocols, one 
%Many authors have proposed definitions to this concept, though most use examples to explain it or adopt a restrictive definition, that may not contemplate a set of systems. <- wrong. Most are wide or use synonyms
%Synthetizing previous definitions, Dey and Abowd have proposed that context is any relevant information used to characterize the situation of entities, specifically: people, places and things.

Context aware (or context sensitive) applications is a designation to those applications that not only are capable of knowing the context of the environment, but also react to it either by means of changing the environment or in a software level. 
% TODO: citation

% [distributed systems]
Ubiquitous systems are a class of distributed systems. So, our approach incorporates well established distributed system concepts that are essential to disseminate and manage context information. Besides this, a distributed component model is adopted to implement the entities that live and interact in the ubiquitous environment.
% FIXME: prooven?

@@

% The general architecture of the framework is expressed in a form of a distributed components model, which defines the RA structure 
The general architecture of the framework includes a Distributed Components Model, which defines the RA structure and the Resource Management Support Services (RMSS). The Figure~\ref{fig:arch} presents a layered architecture that highlights main aspects of the framework. The layers are composed by the following items.

@@

  A resource may be defined as any hardware or software entity that exposes services (of data acquisition or actuation) which can be used by other entities and applications present in IE. Thus, sensors, actuators and smart devices (e.g. stove, fridge, air-conditioning), as well as software modules that provide a service to the environment, fall within this definition.

  The RAs are entities that represent resources. They encapsulate the specificities of resources and expose their information to the environment through standard interfaces, so that others can access them uniformly reducing the complexity of integration.

@@

    \subsubsection{Security Management Service}
    The security concern is always a central topic on UbiComp systems, due to the diverse aspects of security required~\cite{kim2010challenges}. Some aspects are already inherited, depending on the technology adopted, e.g., the \mbox{Wi-Fi} already ensures against the registration of outside devices and the exchange of information. Thus our proposal for the the SecMS discusses requirements at a higher level, and issues at a lower level are defined according to the platform used.

    % [credentials]
    In UbiComp, any human-machine interaction should be as inconspicuous as possible. However, on what matters to the users there is a need for authentication what may happens through explicit interaction, such as entering a password, fingerprint, eye retina recognition, and so on. Other less intrusive ways are progressing, such as the facial recognition, but this raises privacy issues. The users fit in user groups that restricts their access to domains in the IE. For instance, in a smart home the father and the mother may have administrative permissions, below them there are the other occupants (such as elderly and older children) that have some level of permissions, next the younger children, and lastly the guests.

%     The domains are responsible for grouping UbiApps with the same caracte
%     A domain may represent a set of UbiApps with
    A domain can represent a set of applications that share a common characteristic (e.g., home health care, home surveillance, multimedia), or a set of resource within the same constraints (e.g., adult, children, guest). The SecMS allows modifications on security take effect on the level of middleware. Neither a user can manipulate resources or UbiApps that she has not permission to, nor an application can manipulate resources from outside its domain. The specification of domains complies with the theory of sets, thus an entity may be contained in more than one domain, and so on.

@@

  \subsection{Context Interpretation} \label{subsec:ci}
%   [motivation]
  Considering the multitude of resources present in the IE, the context interpretation aims at providing a higher level context information hence allowing the reuse of the logic and a separation of concerns. This feature helps to solve the challenge related to the \textit{variety of context information and services}, what is possible via the RA's interfaces, the CVs and OPs.

@@

  An example of context rules has been presented on Figure~\ref{fig:arch}. At the Applications Layer a set of two components implements the interpretation-actuation pattern: the ``Forgotten Stove Interpreter'' and the ``Reminder Actuator'' respectively, where the goal of the rule is to remind the elderly that he has forgotten the oven on while he went to sleep. The former component, which is a CI, is subscribed to two RAs, namely the ``Bed RA'' and the ``Stove RA'' (the dashed arrow means notify subscribers). The latter is subscribed to the CI and calls the RAs, namely the ``TV RA'' and the ``Clock RA'', directly using RPC (represented by the straight arrow). Internally to the CI a rule expression relates the RAs as  ``bed is occupied and oven is turned on for 30 min'', which means that the CI only will be valid if this condition remains true during ``30 min''. When the CI changes its state from invalid to valid (what includes the timing), the subscribers are notified, the ``Reminder Actuator'' in case. This actuator, in its turn, calls the OPs ``fire alarm'' from ``Clock RA'' and ``set message'' from ``TV RA'', with the parameter ``You forgot the oven on!''. Then it is expected that an accident is prevented.

@@

%   [timer can be CV]
  Another aspect of context rules is the time dimension. The definition of timers inside a rule expression can be done by subscribing the CI to a Timer RA, hence enabling the direct reference in the expression. The timer can also be relative or absolute. An absolute timer is objective: if the timer has reached some timestamp, or if it is within a window of time, then it is valid. %TODO: [+++]

@@

%   [how to aggregate others standards such as rule engines]
  % [rule engine]
  The solution for implementing the context rules within the framework may not be followed strictly as presented. The framework is flexible enough to permit other forms of context interpretation, such as by aggregating a off-the-shelf rule engine (e.g., Jess, Drools, Clips). A rule engine is a tool generally centralized, which receives context information from various sources and processes rules previously inserted, where it is common to use an algorithm such as RETE~\cite{forgy1982rete}, processing rules built exclusively with the clause ``AND'' efficiently. Thus, the developer can use a different approach for rule processing but still enjoying the framework's features for context acquisition, discovery, and so on.
%Removi citação para ~\cite{kaczor2010rule}

@@

  Through the CI the framework address the challenge (ii), which is related to the amount and diversity of context information and services, since the context interpretation allows the autonomous behaviour of RAs.

@@

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid}~\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it can be widely used in sensors and devices embedded in furnitures and home appliances. In addition, this option helped us to implement emulators of several devices using cheap Android cellphones an tablets.

@@

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possibility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do. Within the market, the UbiApp could be tested towards discovering any possible harm that the application could do and any possible inefficiency that could consume too much the energy resources of the host device.
% TODO: talk about market place
% [security concern on store: helps the discovery of malware before the user download]

%   \subsection{Location}
%   
% %   [talk about location]
%   
%   The devices' indoor localization may be acquired through many different approaches, as an example, using radio frequency devices such as RFID tags and Wifi devices~\cite{Talukder2007Tracker} or even a smart floor~\cite{Branzel2013SmartFloor} which can locate people and objects by a pressure-sensing floor. However, the SmartAndroid doesn't implement a specific technique to locate resources, in order to keep the framework generic. The SmartAndroid API deal with location in a hi-level way, making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

  %FIXME: talk about location internally or remove this subsection

  \subsection{Communication Support}
  %[talk about stub intantiation calls]
  %TODO: example of CV and OP? Maybe it is worth add some on the text
  To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. The stubs could also be automatically generated since their role is to forward the calls to the implemented RA. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
  \begin{enumerate}
    \item The UbiApp calls a method from the stub 
    \item The call is forwarded to the RA instance through the network in the second step
    \item This instance can interact with the device (the lamp, in case)
    \item The result or the acknowledge is then returned to the caller
  \end{enumerate}

%     In the first step the UbiApp calls a method from the stub whose call is forwarded to the RA instance throught the network in the sencond step, this instance can interact with the device (the lamp, in case) and return an answer or an aknowledge. In this example, the communication could be used to retrieve come CV status from the lamp such as ``is on'' or ``light intensity''

@@

  The second stage happens when the first RA is already subscribed to the second and a notification routine is triggered, what was represented by the Figure~\ref{fig:pubsub}.b and works as follows:
  \begin{enumerate}
    \item The device has its context changed, what may be caused directly (e.g., a user turning the lamp on) or through the RA (i.e. an application, another RA, or rule, calling some RA's OP which caused the change of context)
    \item Then the RA that represents the device notifies all subscribers to that CV (e.g., ``Is on'' from this Lamp RA)
    \item The subscriber may call some callback method of the UbiApp
  \end{enumerate}

@@

% 
%   \subsection{Distributed Components Model}
% %   [implementation of RA and services]
%   
%   [initialization]
% %   O SmartAndroid é um aplicativo que inicializa o SGAR, repositório e serviços de comunicação. Este aplicativo deve ser executado em uma única unidade do ambiente, deixando o Deamon de Comunicação em estado de espera até receber requisições de outras aplicações. Em outras unidades (clientes) constam os aplicativos instalados com a biblioteca do SmartAndroid, a qual é necessária para executar o serviço de comunicação e \textit{proxys} dos serviços do SGAR. Assim cada unidade pode consultar sobre IARs, através do SDR, obtendo seus ODARs, cada um com o respectivo RANS, o qual é utilizado pelo \textit{proxy} como referência de acesso. 
% 
% 
%   with the RANS 
%   RPC search 
% 
%   
%   [annotation]

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/publish-subscribe-ar-en-v2.pdf}
  \caption{Resource Agent} 
  \label{fig:ra-intercept}
  \end{figure}

  \subsection{Context Rules Composition} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The composition of context rules can be done basically in two ways. The former is at the implementation level, where developers are able define their context rules in a description file or using a programming language. The latter is at end user level, where non trained people are able to define their context rules supported by the PIPA (Subsection~\ref{subsec:ipgap}).

@@

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format, thus the CI code and the binding with the actuators can be automatically generated. The file must contain the IC definition with the rule expression, and may contain the definition of actuators too. This expression contains conditions that are the comparison of RA's CVs with constant values (e.g., Stove.OvenTemperature = $180\,^{\circ}\mathrm{C}$), of two different CVs from the same or from different RAs (e.g., ArConditioning.Temperature $>$ OutsideSensor.Temperature), or the RA's CV alone if it is has a boolean value (e.g., Stove.IsOn). The conditions are bound to each other through the logical operators ``AND'' and ``OR'', it is also allowed the use of ``NOT'' and of parenthesis, what makes possible to compose more complex expressions by using subexpressions. The expression or subexpression may also contain a timer.

  The definition of actuators is simpler than the definition of the CI and is not mandatory, since other actuators can be bound to the IC later. A actuator is composed by a list of actions, where each action have a RANS, a OP reference and the parameters (e.g., Stove.SetOvenTemperature to $100\,^{\circ}\mathrm{C}$). Both the actuator and the CI have a name, that makes easier to search for them.

  The SmartAndroid has a parser that can read the JSON file and create the instances of the CI and actuators. After reading the rule expression the parser creates a tree data structure that corresponds to that expression in memory, and the timers are also referenced within the structure. The CI created queries all the CVs that appear on the expression to keep a cache of the context information and to start the evaluation, next the CI subscribes to all these CVs and sets itself as ``running''. After the creation of the CI, the parser creates instances for the actuators, which are subscribed to the CI previously instantiated.

>>>>

The concepts of the framework were materialized on a platform called \textit{SmartAndroid}~\footnote{For more information visit www.tempo.uff.br/smartandroid}, whose development has enabled appraisal in order to prove that the conceptual framework facilitates the process of building UbiApps. We implemented example applications to validate different aspects of the framework proposed and to testify its capabilities.%, supported by the elaboration of competency questions, which is a mechanism mostly used to evaluate ontologies, but can also be used as an assessment to this work.

The remainder of this paper is organized as follows. First we compare our approach with related works in Section~\ref{sec:related}, then in Section~\ref{sec:overview} we present an overview of the main concepts used as a basis for the development of the framework. In Section~\ref{sec:framework}, we describe the framework's overall architecture as well as, in Section~\ref{sec:features}, its most important features. Our framework reference implementation as the assumptions, techniques and features related are discussed in Section~\ref{sec:smartandroid}. A proof of concept demonstrating the feasibility of building UbiApps using the framework is presented in Section~\ref{sec:evaluation}, showing examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:conclusion} we conclude this paper with the main remarks and future works.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work} \label{sec:related}
% [ubicomp is nice but it is not priceless]
The benefits to users and developers that UbiComp foresees reach beyond the horizon of many imaginative researchers. The possibilities that arise from it surpass the bounds established by common interface devices, such as mice, keyboards and touch screens. Nonetheless, building such adaptable applications requires much effort from developers who see themselves immersed in a universe of device's specifications and communication technologies, apart from the problem itself that the application must solve. Therefore, many researchers have focused on diminishing those obstacles by providing abstractions, middlewares, services, tools, and other supportive techniques, not only for development but also for the prototyping of UbiApps.

@@

%%%%%%%%%%%%%%% CONTEXT AWARENESS
%TODO: Context Toolkit - Dey
%also a famous approach
%we use these abstractions, the aggregation, interpretation, the widget
%This approach provides a starting point
% The Context Toolkit~\cite{dey2001conceptual} is a seminal work that provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation to support rapid prototyping of context aware applications. 
The framework of Dey and Abowd~\cite{dey2001conceptual}, implemented by the Context Toolkit, is a seminal work for the supporting of rapid prototyping of context aware applications and provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation. Another pioneer in the field, the Gaia middleware~\cite{ranganathan2003middleware} was designed to facilitate the construction of applications for IE. It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others. However, these two approaches do not have concerns on the user side and their preferences, including the management and testing of UbiApps at user level.
% Being one pioneer in the field, 
%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets information, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

% DIOS++
The framework proposed in DIOS++~\cite{liu2003dios++} offers abstractions to manipulate distributed sensors and actuators, a communication layer to manage these devices, and a distributed rule engine. But concerns on temporization of context rules and creation of rules at end-user level were not touched upon.

% CASS
The CASS~\cite{Park2007simulation} is a rule simulation environment that have a simple mechanism to discover possible conflicts between rules that are composed only using ``AND'' clause. However, it does not deploy the rules in a running environment.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview} \label{sec:overview}
% [context/contx sensibility]
It is easy to notice that the context information is a first concern topic in ubiquitous systems. But what is ``context''? Many authors have proposed definitions to this concept, though most are little accurate~\cite{baldauf2007survey}. Synthesizing previous definitions, Dey and Abowd have proposed that context is ``any information that can be used to characterize the situation of entities (i.e., whether a person, place or object) that are considered relevant to the interaction between a user and an application, including the user and the application themselves''~\cite{dey2001conceptual}. In a IE ``places'' are the rooms, the floors of a building, or spaces in general that enable the localization of other entities, ``people'' are individuals that populate the environment and interact with it, and ``things'' are virtual representations of physical objects or software components.
%Considering the diversity of devices, systems and protocols, one 
%Many authors have proposed definitions to this concept, though most use examples to explain it or adopt a restrictive definition, that may not contemplate a set of systems. <- wrong. Most are wide or use synonyms
%Synthesizing previous definitions, Dey and Abowd have proposed that context is any relevant information used to characterize the situation of entities, specifically: people, places and things.

Context aware (or context sensitive) applications is a designation to those applications that not only are capable of knowing the context of the environment, but also react to it either by means of changing the environment or in a software level. Thus, the UbiComp is context aware by definition, but the reverse is not necessarily true, since a context aware system may not be designed for IE.
% TODO: citation

% [distributed systems]
Ubiquitous systems are a class of distributed systems. So, our approach incorporates well proven distributed system concepts that are essential to disseminate and manage context information. Besides this, a distributed component model is adopted to implement the entities that live and interact in the ubiquitous environment.
% FIXME: strange

@@

% The general architecture of the framework is expressed in a form of a distributed components model, which defines the RA structure 
The general architecture of the framework includes a Distributed Components Model, which defines the RA structure and the Resource Management Support Services. The Figure~\ref{fig:arch} presents a layered architecture that highlights main aspects of the framework. The layers are composed as follows.

@@

  A resource may be defined as any hardware or software component that exposes services (of data acquisition or actuation) which can be used by other entities and applications present in IE. Thus, sensors, actuators and smart devices (e.g. stove, fridge, air-conditioning), as well as software modules that provide a service to the IE, fall within this definition.

  The RAs are entities that represent resources. They encapsulate the specificities of resources and expose their information to the IE through standard interfaces, so that others can access them uniformly reducing the complexity of integration.

@@

    \subsubsection{Security Management Service}
    The security concern is always a central topic on UbiComp systems, due to the diverse aspects of security required~\cite{kim2010challenges}. Some aspects are already inherited, depending on the technology adopted, e.g., the WPA protocol for \mbox{Wi-Fi} already ensures against the registration of outside devices and the secure exchange of information. %Thus our proposal for the the SecMS discusses requirements at a higher level, and issues at a lower level are defined according to the platform used.

    % [credentials]
    In UbiComp, any human-machine interaction should be as inconspicuous as possible. However, on what matters to end-users there is a need for authentication, what may be realized through explicit interaction, such as entering a password, fingerprint, eye retina recognition, and so on. Besides, other less intrusive ways are progressing, such as the facial recognition, though this raises privacy issues. On our approach, we do not specify which technology is used to guarantee the authentication. But we define user groups that restrict their access to domains in the IE. For instance, in a smart home the father and the mother may have administrative permissions, below them there are the other occupants (such as elderly and older children) that have a lower permission level, followed by the younger children and the guests at last.

%     The domains are responsible for grouping UbiApps with the same caracte
%     A domain may represent a set of UbiApps with
    A domain can represent a set of applications that share a common characteristic (e.g., home health care, home surveillance, multimedia), or a set of resource with the same constraints (e.g., adult, children, guest). Neither a user can manipulate resources or UbiApps that he has not permission to, nor an application can manipulate resources from outside its domain. The specification of domains complies with the theory of sets, thus an entity may be contained in more than one domain, and so on.
%     The SecMS allows modifications on security to take effect on the level of the middleware.

@@

  \subsection{Context Interpretation} \label{subsec:ci}
%   [motivation]
  Considering the multitude of resources present in the IE, the context interpretation aims at providing a higher level context information hence allowing the reuse of the logic and the separation of concerns. This feature helps to solve the challenge \textit{variety of context information and services} (ii), what is possible via the RA's interfaces, the CVs and OPs.

@@

  An example of context rules has been presented on Figure~\ref{fig:arch}. At the Applications Layer a set of two components implements the interpretation-actuation pattern: the ``Forgotten Stove Interpreter'' and the ``Reminder Actuator'' respectively, where the goal of the rule is to remind the elderly that he has forgotten the oven on while he went to sleep. The former component, which is a CI, is subscribed to two RAs, namely the ``Bed RA'' and the ``Stove RA'' (the dashed arrow means ``notify subscribers''). The latter is subscribed to the CI and calls the RAs, namely the ``TV RA'' and the ``Clock RA'', directly using RPC (represented by the straight arrow). Internally to the CI a rule expression relates the RAs as  ``bed is occupied and oven is turned on, for 30 min'', which means that the CI only will be valid if this condition remains true during ``30 min''. When the CI changes its state from invalid to valid (what includes the timing), the subscribers are notified, the ``Reminder Actuator'' in case. This actuator, in its turn, calls the OPs ``fire alarm'' from ``Clock RA'' and ``set message'' from ``TV RA'', with the parameter ``You forgot the oven on!''. Then it is expected that an accident is prevented.

@@

%   [timer can be CV]
  Another aspect of context rules is the time dimension. The definition of timers inside a rule expression can be done by subscribing the CI to a Timer RA, hence enabling the direct reference in the expression. The timer can also be relative or absolute. The absolute timer is objective: if the timer has reached some timestamp (e.g., 14:00), or if it is within a window of time (e.g., above 12:00 and below 14:00), then it is valid. The relative timer is more common, it counts the time from now until the specified timestamp (e.g., 30 min from now).

@@

%   [how to aggregate others standards such as rule engines]
  % [rule engine]
  The solution for implementing the context rules within the framework may not be followed strictly as presented. The framework is flexible enough to permit other forms of context interpretation, such as by aggregating a off-the-shelf rule engine (e.g., Jess, Drools, Clips). A rule engine is a tool generally centralized, which receives context information from various sources and processes rules previously inserted, where it is common to use an algorithm such as RETE~\cite{forgy1982rete}, that processes rules built exclusively with the clause ``AND'', but very fast. Thus, the developer can use a different approach for rule processing but still enjoying the framework's features for context acquisition, discovery, and so on.
%Removi citação para ~\cite{kaczor2010rule}

@@

%   Through the CI the framework address the challenge (ii), which is related to the amount and diversity of context information and services, since the context interpretation allows the autonomous behaviour of RAs.

@@

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{\textit{SmartAndroid}}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The \textit{SmartAndroid} has been developed over Google Android platform as we believe that in a near future it can be widely used in sensors and devices embedded in furnitures and home appliances. In addition, this option helped us to implement emulators of several devices using cheap Android cellphones an tablets.

@@

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The \textit{SmartAndroid} middleware contemplates the possibility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a \textit{SmartAndroid} Market, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do. Within the market, the UbiApp could be tested towards discovering any possible harm that the application could do and any possible inefficiency that could consume too much the energy resources of the host device.
% ~\cite{google2013}
% [security concern on store: helps the discovery of malware before the user download]

%   \subsection{Location}
%   
% %   [talk about location]
%   
%   The devices' indoor localization may be acquired through many different approaches, as an example, using radio frequency devices such as RFID tags and Wifi devices~\cite{Talukder2007Tracker} or even a smart floor~\cite{Branzel2013SmartFloor} which can locate people and objects by a pressure-sensing floor. However, the \textit{SmartAndroid} doesn't implement a specific technique to locate resources, in order to keep the framework generic. The \textit{SmartAndroid} API deal with location in a hi-level way, making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

  \subsection{Communication Support}
  %[talk about stub intantiation calls]
  %TODO: example of CV and OP? Maybe it is worth add some on the text
  To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. The stubs could also be automatically generated since their role is to forward the calls to the implemented RA. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
  \begin{enumerate}
    \item The UbiApp calls a method from the stub 
    \item The call is forwarded to the RA instance through the network in the second step
    \item This instance interacts with the device (i.e. the lamp)
    \item The result or the acknowledge is then returned to the caller
  \end{enumerate}
  %FIXME gain line: can remove work acknowledge

@@

  The second stage happens when the first RA is already subscribed to the second and a notification routine is triggered, what was represented by the Figure~\ref{fig:pubsub}.b and works as follows:
  \begin{enumerate}
    \item The device has its context changed, what may be caused directly (e.g., a person lights the lamp) or via the RA (i.e. an application, RA, or rule calls some RA's OP causing the context change)
    \item Then the device's RA notifies all subscribers of that CV (e.g., ``Is on'' from this Lamp RA)
    \item The subscriber calls a callback method of the UbiApp
  \end{enumerate}

@@

% 
%   \subsection{Distributed Components Model}
% %   [implementation of RA and services]
%   
%   [initialization]
% %   O \textit{SmartAndroid} é um aplicativo que inicializa o SGAR, repositório e serviços de comunicação. Este aplicativo deve ser executado em uma única unidade do ambiente, deixando o Deamon de Comunicação em estado de espera até receber requisições de outras aplicações. Em outras unidades (clientes) constam os aplicativos instalados com a biblioteca do \textit{SmartAndroid}, a qual é necessária para executar o serviço de comunicação e \textit{proxys} dos serviços do SGAR. Assim cada unidade pode consultar sobre IARs, através do SDR, obtendo seus ODARs, cada um com o respectivo RANS, o qual é utilizado pelo \textit{proxy} como referência de acesso. 
% 
% 
%   with the RANS 
%   RPC search 
% 
%   
%   [annotation]

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3in]{figures/publish-subscribe-ar-en-v3.pdf}
  \caption{Resource Agent} 
  \label{fig:ra-intercept}
  \end{figure}

  \subsection{Context Rules Composition} \label{subsec:rules}
%   [implementation of context rules]
  The \textit{SmartAndroid} supports the context interpretation by providing mechanisms to create context rules. The composition of context rules can be done basically in two ways. The former is at the implementation level, where developers are able define their context rules in a description file or using a programming language. The latter is at end-user level, where non trained people are able to define their context rules supported by the PIPA (Subsection~\ref{subsec:ipgap}).

@@

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format, thus the CI code and the binding with the actuators can be automatically generated. The file must contain the CI definition with the rule expression, and may contain the definition of actuators too. This expression contains conditions that are the comparison of RA's CVs with constant values (e.g., Stove.OvenTemperature = $180\,^{\circ}\mathrm{C}$), of two different CVs from the same or from different RAs (e.g., ArConditioning.Temperature $>$ OutsideSensor.Temperature), or the RA's CV alone if it has a boolean value (e.g., Stove.IsOn). The conditions are bound to each other through the logical operators ``AND'' and ``OR'', it is also allowed the use of ``NOT'' and of parenthesis, what makes possible to compose more complex expressions by using subexpressions. The expression or subexpression may also contain a timer. 
  %TODO: explain why to use subexpressions (example: person at bed or at sofa)

  The definition of actuators is simpler than the definition of the CI and is not mandatory, since other actuators can be bound to the CI later. A actuator is composed by a list of actions, where each action have a RANS, a OP reference and the parameters (e.g., Stove.SetOvenTemperature to $100\,^{\circ}\mathrm{C}$). Both the actuator and the CI have a name, what makes it easier to search for them later.

  The \textit{SmartAndroid} has a parser that can read the JSON file and generates the instances of the CI and actuators. After reading the rule expression the parser creates a tree data structure that corresponds to that expression in memory, and the timers are also referenced within the structure. The CI queries all the CVs that appear on the expression to keep a cache of the context information and to start the evaluation. Next the CI subscribes to all these CVs and sets itself as ``running''. After the creation of the CI, the parser creates instances for the actuators, which are subscribed to the CI previously instantiated.
